# Additional analysis

> From: Carin Waslo   
> Sent: Wednesday, August 08, 2018 3:27 PM  
> To: Benjamin Chan <chanb@ohsu.edu>  
> Subject: Spain - Glutathione ratios as the mechanism of action of LA in PMS  
> 
> Good afternoon,
> 
> Dr. Spain and I are developing a poster and abstract, for the Glutathione study,
> to submit the ACRIMS 2019. Would you be willing to 1) review the attached
> abstract and poster to ensure correct interpretation of the results you
> provided, and 2) run one last analysis to see if baseline GSH:GSSG ratios
> "predict" subsequent (at month 24) % change brain volumes for LA and placebo
> groups?
> 
> Please let me know if you have any questions or need more information. And thank
> you very much!
> 
> Best,
> Carin

The model closely follows the model used for Aim 2.

The model is

$$
y_{\text{% brain atrophy}, i} =
\beta_0
+ \beta_1 x_{\text{Baseline GSH:GSSG}, i}
+ e_{i}
$$

Linear model.

```{r}
M <- lm(brainAtrophy ~ concRatioM0, data = df2)
summary(M)
M1 <- update(M, ~ . + studyArm)
summary(M1)
M2 <- update(M1, ~ . + studyArm * concRatioM0)
summary(M2)
tidy(M2) %>% kable(digits = c(2, 2, 2, 2, 4))
```

Details (not shown).

```{r additionalAnalysisDiagnostics, results = "hide", fig.show = "hide"}
res <- residuals(M)
plot(res)
qqnorm(res)
plot(M)
```


## Interpretation

```{r, results = "hide"}
partcor <- 
  df2 %>% 
  dplyr::select(brainAtrophy, concRatioM0) %>% 
  as.matrix() %>% 
  pcor()
r <- partcor %>% .$estimate %>% .[1, 2]
rsq <- r^2 
```

* Percent brain atrophy was not significantly associated with baseline GSH:GSSG concentration
  * $\hat{\beta}$ of `r M %>% tidy() %>% filter(term == "concRatioM0") %>% pull(estimate) %>% signif(3)` (p-value = `r M %>% tidy() %>% filter(term == "concRatioM0") %>% pull(p.value) %>% signif(3)`)
  * Correlation coefficient = `r r %>% signif(3)`
  * Baseline GSH:GSSG concentration explains `r sprintf("%.3g%%", rsq * 100)` of the variation in brain atrophy
```{r, results = "hide"}
partcor <- 
  df2 %>% 
  select(brainAtrophy, concRatioM0, studyArm) %>%
  mutate(studyArm = as.integer(unclass(studyArm))) %>% 
  as.matrix() %>% 
  pcor()
r <- partcor %>% .$estimate %>% .[1, 2]
rsq <- r^2 
```
* The association was still not significant after adjusting for study arm
  * $\hat{\beta}$ of `r M1 %>% tidy() %>% filter(term == "concRatioM0") %>% pull(estimate) %>% signif(3)` (p-value = `r M1 %>% tidy() %>% filter(term == "concRatioM0") %>% pull(p.value) %>% signif(3)`)
  * Partial correlation coefficient = `r r %>% signif(3)`
  * Baseline GSH:GSSG concentration explains `r sprintf("%.3g%%", rsq * 100)` of the variation in brain atrophy after adjusting for study arm
