# Import raw data

```{r}
df1 <-
  "data/raw/Spain Manifest.xlsx" %>% 
  read_excel(.) %>% 
  rename(sampleID = "Sample #",
         patientID = "Patient ID",
         visitMonth = "Visit month",
         studyArm = "Study Arm",
         brainAtrophy = "M24 brain atrophy (% change)") %>% 
  mutate(patientID = patientID %>% factor(),
         visitMonth = visitMonth %>% gsub("M", "", .) %>% as.integer() %>% factor(),
         studyArm = studyArm %>% factor())
```

```{r}
foo <- function(range) {
  require(magrittr)
  require(readxl)
  "data/raw/Copy of Oregon GSHGSSG.xlsx" %>% 
  read_excel(., range = range)
}
df2 <-
  foo("G2:I60") %>% 
  rename(sampleID = "X__1",
         meanResponseGSH = "GSH",
         meanResponseGSSG = "GSSG")
df3 <-
  foo("K2:M60") %>% 
  rename(sampleID = "X__1",
         sdGSH = "GSH",
         sdGSSG = "GSSG")
df4 <-
  foo("O2:Q60") %>% 
  rename(sampleID = "X__1",
         seGSH = "GSH",
         seGSSG = "GSSG")
df5 <-
  foo("S2:W60") %>% 
  rename(sampleID = "X__1",
         concentrationGSHugml = "GSH",
         concentrationGSSGugml = "GSSG",
         concentrationStdEGSHugml = "GSH StdE",
         concentrationStdEGSSGugml = "GSSG StdE")
df6 <-
  foo("Y2:AC60") %>% 
  rename(sampleID = "X__1",
         concentrationGSHuM = "GSH",
         concentrationGSSGuM = "GSSG",
         concentrationStdEGSHuM = "GSH StdE",
         concentrationStdEGSSGuM = "GSSG StdE")
df7 <-
  foo("G64:K122") %>% 
  rename(sampleID = "X__1",
         nmolGSH = "GSH",
         nmolGSSG = "GSSG",
         nmolStdEGSH = "GSH StdE",
         nmolStdEGSSG = "GSSG StdE")
df8 <-
  foo("M64:P122") %>% 
  rename(sampleID = "X__1",
         proteinConcugul = "(ug/ul)",
         proteinConcug = "ug",
         proteinConcmg = "mg")
df9 <-
  foo("R64:V122") %>% 
  rename(sampleID = "X__1",
         normalizedConcGSH = "GSH",
         normalizedConcGSSG = "GSSG",
         normalizedConcStdEGSH = "GSH StdE",
         normalizedConcStdEGSSG = "GSSG StdE")
df10 <-
  foo("X64:Y122") %>% 
  rename(sampleID = "X__1",
         normalizedConcRatio = "GSH/GSSG")
```

Joined data set.
Include only the 

* Study design variables,
* Normalized concentrations, and 
* GSH/GSSG ratios

```{r}
df <-
  inner_join(df1, df9 , by = "sampleID") %>% 
  inner_join(.  , df10, by = "sampleID")
```

```{r}
rm(df1, df2, df3, df4, df5, df6, df7, df8, df9, df10, foo)
```


## Clean data

Data check **before cleaning**.

```{r}
df %>% 
  group_by(patientID) %>% 
  summarize(hasM0 = sum(visitMonth == 0),
            hasM3 = sum(visitMonth == 3),
            hasM12 = sum(visitMonth == 12)) %>% 
  filter(hasM0 != 1 | hasM3 != 1 | hasM12 != 1) %>% 
  kable(align = "c")
```

Recode data per Cassidy's email

> From: Cassidy Taylor  
> Sent: Monday, March 12, 2018 3:17 PM  
> To: Benjamin Chan <chanb@ohsu.edu>; Rebecca Spain <spainr@ohsu.edu>  
> Subject: RE: Rebecca Spain Glutathione Study  
> 
> Just heard back from the lab and based on the remaining levels of blood, the
> manifest was incorrectly labeled. It should read
> 
> Sample 44 = 143 M0  
> Sample 45 = 143 M3  
> Sample 46 = 143 M12  

```{r}
df[df$sampleID == 44, "visitMonth"] <- 0
df[df$sampleID == 46, "patientID"] <- 143
df[df$sampleID == 46, "studyArm"] <- "Placebo"
```

Data check **after cleaning**.
Should return a data frame with only `patientID %in% c(120, 122)`.

```{r}
df %>% 
  group_by(patientID) %>% 
  summarize(hasM0 = sum(visitMonth == 0),
            hasM3 = sum(visitMonth == 3),
            hasM12 = sum(visitMonth == 12)) %>% 
  filter(hasM0 != 1 | hasM3 != 1 | hasM12 != 1) %>% 
  kable(align = "c")
```

Double-check data from `patientID == 143`.

```{r}
df %>% 
  filter(patientID == 143) %>% 
  kable()
```


## Create analytic data frames for each aim

Aim 1

```{r}
df1 <- 
  df %>% 
  select(-brainAtrophy)
str(df1)
```

Aim 2

```{r}
df2 <- 
  df %>% 
  filter(visitMonth %in% c(0, 12)) %>% 
  select(patientID, studyArm, visitMonth, normalizedConcRatio) %>% 
  mutate(visitMonth = sprintf("concRatioM%s", as.character(visitMonth))) %>% 
  spread(visitMonth, normalizedConcRatio) %>% 
  mutate(pctChangeConcRatio = (concRatioM12 - concRatioM0) / concRatioM0 * 100) %>% 
  inner_join(df %>% select(patientID, brainAtrophy) %>% filter(!is.na(brainAtrophy)),
             by = "patientID")
str(df2)
df2 %>% arrange(studyArm, patientID) %>% kable(digits = 2)
```
